### 2020-12-09 tbl_tbl_summary_args.tex
in progress

Table: `tbl_summary()` function arguments

|Argument       |Description                                                      |
|:--------------|:----------------------------------------------------------------|
|`label`        |specify the variable labels printed in table                     |
|`type`         |specify the variable type (e.g. continuous, categorical, etc.)   |
|`statistic`    |change the summary statistics presented                          |
|`digits`       |number of digits the summary statistics will be rounded to       |
|`missing`      |whether to display a row with the number of missing observations |
|`missing_text` |text label for the missing number row                            |
|`sort`         |change the sorting of categorical levels by frequency            |
|`percent`      |print column, row, or cell percentages                           |