### 2020-12-09 tbl_trial_desc.tex

\begin{table}

\caption{\label{tab:unnamed-chunk-1}Example data frame, `trial`}
\centering
\begin{tabular}[t]{l|l|l|l}
\hline
colname & label & class & values\\
\hline
trt & Chemotherapy Treatment & character & ` Drug A `, ` Drug B `\\
\hline
age & Age & numeric & ` 6 `, ` 9 `, ` 10 `, ` 17 ` , ...\\
\hline
marker & Marker Level (ng/mL) & numeric & ` 0.003 `, ` 0.005 `, ` 0.013 `, ` 0.015 ` , ...\\
\hline
stage & T Stage & factor & ` T1 `, ` T2 `, ` T3 `, ` T4 `\\
\hline
grade & Grade & factor & ` I `, ` II `, ` III `\\
\hline
response & Tumor Response & integer & ` 0 `, ` 1 `\\
\hline
death & Patient Died & integer & ` 0 `, ` 1 `\\
\hline
ttdeath & Months to Death/Censor & numeric & ` 3.53 `, ` 5.33 `, ` 6.32 `, ` 7.27 ` , ...\\
\hline
\end{tabular}
\end{table}