### 2020-12-09 tbl_tbl_summary_args.tex
in progress

\captionsetup[table]{labelformat=empty,skip=1pt}
\begin{longtable}{ll}
\toprule
Argument & Description \\ 
\midrule
\texttt{label} & specify the variable labels printed in table \\ 
\texttt{type} & specify the variable type (e.g. continuous, categorical, etc.) \\ 
\texttt{statistic} & change the summary statistics presented \\ 
\texttt{digits} & number of digits the summary statistics will be rounded to \\ 
\texttt{missing} & whether to display a row with the number of missing observations \\ 
\texttt{missing\_text} & text label for the missing number row \\ 
\texttt{sort} & change the sorting of categorical levels by frequency \\ 
\texttt{percent} & print column, row, or cell percentages \\ 
\bottomrule
\end{longtable}