### 2020-12-09 tbl_trial_desc.tex

\captionsetup[table]{labelformat=empty,skip=1pt}
\begin{longtable}{clll}
\toprule
colname & label & class & values \\ 
\midrule
\texttt{\textquotedbl{}trt\textquotedbl{}} & Chemotherapy Treatment & character & \texttt{Drug A}, \texttt{Drug B} \\ 
\texttt{\textquotedbl{}age\textquotedbl{}} & Age & numeric & \texttt{6}, \texttt{9}, \texttt{10}, \texttt{17} , ... \\ 
\texttt{\textquotedbl{}marker\textquotedbl{}} & Marker Level (ng/mL) & numeric & \texttt{0.003}, \texttt{0.005}, \texttt{0.013}, \texttt{0.015} , ... \\ 
\texttt{\textquotedbl{}stage\textquotedbl{}} & T Stage & factor & \texttt{T1}, \texttt{T2}, \texttt{T3}, \texttt{T4} \\ 
\texttt{\textquotedbl{}grade\textquotedbl{}} & Grade & factor & \texttt{I}, \texttt{II}, \texttt{III} \\ 
\texttt{\textquotedbl{}response\textquotedbl{}} & Tumor Response & integer & \texttt{0}, \texttt{1} \\ 
\texttt{\textquotedbl{}death\textquotedbl{}} & Patient Died & integer & \texttt{0}, \texttt{1} \\ 
\texttt{\textquotedbl{}ttdeath\textquotedbl{}} & Months to Death/Censor & numeric & \texttt{3.53}, \texttt{5.33}, \texttt{6.32}, \texttt{7.27} , ... \\ 
\bottomrule
\end{longtable}