### 2021-04-10 reviewer_responses/gtsummary_review_2021-03-23.Rmd
updates

---
title: "gtsummary Reviewer Responses"
author: "Daniel Sjoberg, Karissa Whiting, Mike Curry"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: sandstone
    toc: yes
    toc_float: yes
    number_sections: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(gtsummary)
library(gt)

```

Thank you for your reviews and for your consideration of our paper. We believe we have adequately addressed the reviewer's issues and have responded to the comments below.

# Reviewer 1

1. The authors didn’t provide any background review. Generating summary stats tables is such a common task that many people have tried to solve this problem before. I hope the authors could provide a background review of at least a few of the following packages

> We agree that a section in the paper mentioning capabilities of other similar table making packages will help contextualize gtsummary's value. We have added this in the revised paper.

2. It seems that gtsummary has very limited support on generating LaTeX tables....It is okay to have limited support on some features but in terms of article writing, I would like to see the author mention it as a limitation in the discussion section.

> The original submission did not make reference specically to the availability of LaTeX output. We've added text to the *Print Engine* section of the manuscript explicicty listing the output options, including LaTeX.

3. The package is well written with solid documentation. I only have one question here. In `tbl_summary`, why let it have its own `by` argument instead of using `dplyr`’s `group_by`?

> One important difference between `dplyr`’s `group_by` and `tbl_summary`'s `by` argument is that you can group by many variables using `group_by`, but only one variable using `by`. Since `tbl_summary()` cannot support all the uses of `group_by`, we wanted to make a clear distinction that these have different functionalities.

# Reviewer 2

1. The article is very heavy in tables, which is understandable given the nature of the package. But maybe there is a way to rationalize? For instance, the `tbl_svysummary_1` table is essentially identical to the previous one. Do we need to see it in such a short exposition, or is it sufficient to discuss survey support?

> We have discovered many of gtsummary's most active users require support for survey data, so would like to keep this example in, unless the editor feels otherwise.

2. I’m not sure I understand the motivation for the table customization functions (e.g., bold_levels). Isn’t the whole point of building a package on top of {gt} that we can leverage its functions directly?

> The reviewer raises a great point. One of the great features of gtsummary is teh accessabitily of every gt function to produce a highly customized table, when needed. We have included functions to assist in common formatting to make them more accessible to users. For example, to re-created the functionality of the `bold_labels()` function, the user would need to pass `as_gt() %>% gt::tab_style(style = gt::cell_text(weight = "bold"), locations = gt::cells_body(columns = gt::vars(label), rows = row_type == "label"))`. The helper function is much easier to read, and moreover, the gt version requires knowledge of hidden columns (e.g. `"row_type"`) to correctly select the label rows. Without these formatting helper functions, most passive users of the package would not have the knowledge to create the formatting on their own.

3. Package Testing

> As noted by the reviewer, many of the unit tests check for specific error/warnings/messages, which in some cases, may not be considered best practices. The {gtsummary} package, however, currently has 1100 unit tests with hundreds of tests that check for quite specific output. The sum total of the tests, in our opinion, provides great unit testing coverege testing a wide variety of options possible in the package.

4. tbl_summary: the continuous/categorical cutoff seems to be 10. can this be changed?

> Although you can't currently adjust the default cutoff, tidyselect helper functions and the `type=`  argument make it very easy to specify what type you want your variables to be.

5. tbl_summary: Is there a way to set the number of digits for all variables at the same time? This can be onerous when there are many variables.

> tbl_summary() works with all tidyselect functions (in addition to tidyselect-like functions exported by {gtsummary}), and you can specify all variables using the `everything()` selector. For example,

```{r}
trial %>% 
  select(age, marker, ttdeath) %>%
  tbl_summary(digits = everything() ~ 2)
```

6. tbl_summary: Can we give a manual sort order?

> The ordering of the variable in `tbl_summary()` aligns with the order the variables appear in the input data frame. To re-arrange the variables, use `select()` before passing the data frame to `tbl_summary()`, e.g. `trial %>% select(age, grade) %>% tbl_summary()`

7. tbl_summary: Is it possible to use custom/arbitrary functions in the statistic argument, or are we limited to those explicitly defined by the package?

> Yes! Any function in base R, another package, or user-defined can be used to summarize the continuos data in `tbl_summary()`. The only restriction is the first argument of the function must accept a vector, and return a scalar.

8. tbl_regression: How do I manually reorder terms?

> The terms appear in the same order as regression function. The simplest way to reorder the terms to order them in, for example, the `lm()` call. If for some reason that is not possible, they can be re-ordered after the `tbl_regression()` table has been built using `modify_table_body()` with `dplyr::arrange()`.